---
title: "OHDSI Evidence Network Summary"
date: "2025-03-12"
format: 
  dashboard
---

This document provides a summary of the OHDSI Evidence Network data sources. 

```{r setup, include=FALSE}
library(tidyverse)
library(htmltools)
library(leaflet)
library(plotly)
library(DT)
library(ggplot2)
library(DatabaseConnector)
library(SqlRender)
library(treemap)
library(igraph)
library(dplyr)
library(scales)     # for label formatting
library(viridis)

# Create variables
oenDatabaseSchema <- "ohdsi_evidence_network"

# Create connection details
dbProfileConnectionDetails <- DatabaseConnector::createConnectionDetails(dbms = "postgresql",
                                                                         user = Sys.getenv("OEN_USER"),
                                                                         password = Sys.getenv("OEN_PASSWORD"),
                                                                         server = Sys.getenv("OEN_SERVER"),
                                                                         port = 5432,
                                                                         extraSettings = "ssl=true&sslfactory=org.postgresql.ssl.NonValidatingFactory")
```


```{r echo=FALSE}
## Get total number of databases
sql <- "Select 'Number of Databases' as statistic, 
                count(*) as value 
        from @oen_database_schema.metadata"
renderedSql <- SqlRender::render(sql, oen_database_schema = oenDatabaseSchema)

conn <- DatabaseConnector::connect(dbProfileConnectionDetails)
totalDbs <- DatabaseConnector::querySql(conn, renderedSql)
DatabaseConnector::disconnect(conn)

totalDbsValue <- totalDbs$VALUE[1]

# totalDbsTile <- flexdashboard::valueBox(
#   value = totalDbsValue,
#   icon = "fa-database",
#   color = "primary"
# )
# 
# totalDbsTile

```


```{r echo = FALSE}
## Get total number of institutions
sql <- "Select count(distinct substring(oen_id, 1, 4)) as total_institutions 
  from @oen_database_schema.metadata"
renderedSql <- SqlRender::render(sql, oen_database_schema = oenDatabaseSchema)

conn <- DatabaseConnector::connect(dbProfileConnectionDetails)
totalInstitutions <- DatabaseConnector::querySql(conn, renderedSql)
DatabaseConnector::disconnect(conn)

totalInstitutionsValue <- totalInstitutions$TOTAL_INSTITUTIONS[1]

# totalInsTile <- flexdashboard::valueBox(
#   value = totalInstitutionsValue,
#   icon = "fa-hospital",
#   color = "info"
# )

#totalInsTile
```

## Row {height=10%}
### Column

```{r}
#| content: valuebox
#| title: "Total Databases"
list(
  icon = "database-fill",
  color = "success",
  value = totalDbsValue
)
```

### Column 
```{r}
#| content: valuebox
#| title: "Total Institutions"
list(
  icon = "hospital-fill",
  color = "info",
  value = totalInstitutionsValue
)
```

## Databases Overview {.tabset}

### Databases by Data Type

```{r echo=FALSE}


## Get databases by data type
sql <- "Select data_type, count(*) as db_count
            from(select
                    case when data_provenance is NULL then 'Other'
                    when data_provenance like 'NA%' then 'Other'
                    else data_provenance end as data_type
                 from @oen_database_schema.metadata
                ) A
        group by data_type"
renderedSql <- SqlRender::render(sql, oen_database_schema = oenDatabaseSchema)

conn <- DatabaseConnector::connect(dbProfileConnectionDetails)
dbDataTypes <- DatabaseConnector::querySql(conn, renderedSql)
DatabaseConnector::disconnect(conn)

dbDataTypes <- dbDataTypes %>%
  mutate(DATA_TYPE = factor(DATA_TYPE, levels = DATA_TYPE[order(-DB_COUNT)]))

  dbDataTypesPlot <- plot_ly(
    dbDataTypes,
    type = "treemap",
    labels = ~DATA_TYPE,
    values = ~DB_COUNT,
    parents = NA,
    textinfo = "label+value+percent entry",
    textposition = "middle center",
    textfont = list(size = 16),  # Increase font size
    marker = list(colors = viridis::viridis(n = nrow(dbDataTypes)))
  ) %>%
    layout(
      title = "Databases by Data Type",
      uniformtext = list(minsize = 20, mode = "show")  # Ensure text is centered and uniform
    )
  dbDataTypesPlot

```



```{r echo=FALSE}
### Geographic Distribution

# leaflet(ohdsi_sites) %>%
#   addTiles() %>%
#   addCircleMarkers(~Lon, ~Lat, popup = ~paste0("<b>", cdm_source_abbreviation, "</b><br>"#,
#                                               # "Country: ", Country, "<br>",
#                                               # "Data Years: ", Data_Years, "<br>",
#                                               # "Patients: ", Patient_Count)
#                                               ))
```

## Demographics Overview {.tabset}

### Gender Distribution

```{r echo=FALSE}

## Get total number of databases
sql <- "Select 'Number of Databases' as statistic, 
                count(*) as value 
        from @oen_database_schema.metadata"
renderedSql <- SqlRender::render(sql, oen_database_schema = oenDatabaseSchema)

conn <- DatabaseConnector::connect(dbProfileConnectionDetails)
totalDbs <- DatabaseConnector::querySql(conn, renderedSql)

## Get databases gender
sql <- "select oen_id, gender_concept_id, concept_name, sum(count_value) as total_persons
        from(
              select oen_id, case when stratum_1 in ('4214687','8551') then '0' else stratum_1 end as gender_concept_id, count_value
            	from @oen_database_schema.db_profiles
            	where analysis_id = 2
        	) a	
        join vocabulary.concept c
	          on cast(a.gender_concept_id as integer) = c.concept_id
        group by oen_id, gender_concept_id, concept_name;"

renderedSql <- SqlRender::render(sql, oen_database_schema = oenDatabaseSchema)

dbGenders <- DatabaseConnector::querySql(conn, renderedSql)
names(dbGenders) <- c("OEN Id", "Gender Concept Id", "Gender", "Count")

DatabaseConnector::disconnect(conn)

demo_plot <- dbGenders %>%
  group_by(Gender) %>%
  summarize(Count = sum(Count), Databases = n_distinct(`OEN Id`)) %>%
  ggplot(aes(x = Gender, y = Count, fill = Databases)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Gender Distribution Across the Network", x = "Gender", y = "Patient Count", fill = "Number of Databases") +
  theme_minimal()

plotly::ggplotly(demo_plot)

```

## Row

## Year of Birth {.tabset}

### Year of Birth Distribution

```{r echo=FALSE}

## Get year of birth by database
sql <- "select oen_id, stratum_1, count_value
        from @oen_database_schema.db_profiles
        where analysis_id = 3 and cast(stratum_1 as integer) > 1889
        order by stratum_1;"

renderedSql <- SqlRender::render(sql, oen_database_schema = oenDatabaseSchema)
conn <- DatabaseConnector::connect(dbProfileConnectionDetails)

dbYearsOfBirth <- DatabaseConnector::querySql(conn, renderedSql)
names(dbYearsOfBirth) <- c("OEN Id", "Year of Birth", "Count")

DatabaseConnector::disconnect(conn)

dbYearsOfBirthSum <- dbYearsOfBirth %>%
  group_by(`Year of Birth`) %>%
  summarize(Count = sum(Count))

yob_plot <- dbYearsOfBirthSum %>%
  ggplot(aes(x = `Year of Birth`, y = Count, fill = `Year of Birth`)) +
  geom_col(position = "dodge") +
  scale_x_discrete(breaks = seq(min(dbYearsOfBirthSum$`Year of Birth`), max(dbYearsOfBirthSum$`Year of Birth`), by = 5)) +
  labs(title = "Year of Birth Distribution Across the Network", x = "Year of Birth", y = "Patient Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  scale_fill_viridis_d()

plotly::ggplotly(yob_plot)

```

## Race and Ethnicity Distribution {.tabset}

### Race Distribution

```{r echo=FALSE}
## Race and Ethnicity {.tabset}
sql <-  "select oen_id, stratum_1 as race_concept_id, concept_name, sum(count_value) as total_persons
    from(
        select oen_id, stratum_1, count_value
    	from @oen_database_schema.db_profiles
    	where analysis_id = 4
    	) a	
    join vocabulary.concept c
    	on cast(a.stratum_1 as integer) = c.concept_id
    group by oen_id, stratum_1, concept_name"
    
    renderedSql <- SqlRender::render(sql, oen_database_schema = oenDatabaseSchema)
    conn <- DatabaseConnector::connect(dbProfileConnectionDetails)
    
    dbRaces <- DatabaseConnector::querySql(conn, renderedSql)


names(dbRaces) <- c("OEN Id", "Race Concept Id", "Race", "Count")

DatabaseConnector::disconnect(conn)

race_plot <- dbRaces %>%
  group_by(Race) %>%
  summarize(Count = sum(Count), Databases = n_distinct(`OEN Id`)) %>%
  ggplot(aes(x = reorder(Race, Databases), y = Databases, fill = Race)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
  labs(title = "Race Distribution Across the Network", x = "Race", y = "Number of Databases") +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_fill_viridis_d()


plotly::ggplotly(race_plot)

```

### Ethnicity Distribution

```{r echo=FALSE}

sql <-  "select oen_id, stratum_1 as ethnicity_concept_id, concept_name, sum(count_value) as total_persons
    from(
        select oen_id, stratum_1, count_value
    	from @oen_database_schema.db_profiles
    	where analysis_id = 5
    	) a	
    join vocabulary.concept c
    	on cast(a.stratum_1 as integer) = c.concept_id
    group by oen_id, stratum_1, concept_name"
    
    renderedSql <- SqlRender::render(sql, oen_database_schema = oenDatabaseSchema)
    conn <- DatabaseConnector::connect(dbProfileConnectionDetails)
    
    dbEthnicities <- DatabaseConnector::querySql(conn, renderedSql)

DatabaseConnector::disconnect(conn)

names(dbEthnicities) <- c("OEN Id", "Ethnicity Concept Id", "Ethnicity", "Count")

ethnicity_plot <- dbEthnicities %>%
  group_by(Ethnicity) %>%
  summarize(Count = sum(Count), Databases = n_distinct(`OEN Id`)) %>%
  ggplot(aes(x = reorder(Ethnicity, Count), y = Databases, fill = Ethnicity)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
  labs(title = "Ethnicity Distribution Across the Network", x = "Ethnicity", y = "Number of Databases") +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_fill_viridis_d()


plotly::ggplotly(ethnicity_plot)

```


## Clinical Context {.tabset}

### Visit Representation

```{r echo=FALSE} 

sql <-  "select dp.oen_id,max(count_value) as person_count, a.total_person_count, 
		max(count_value) / a.total_person_count as prop,
		visit_ancestor_concept_name
          from @oen_database_schema.db_profiles dp 
          join (
          	select oen_id,count_value as total_person_count
          	from @oen_database_schema.db_profiles dp 
            where analysis_id = 1
          )a
          on dp.oen_id = a.oen_id
where analysis_id = 200
group by dp.oen_id, visit_ancestor_concept_name, a.total_person_count"
    
    renderedSql <- SqlRender::render(sql, oen_database_schema = oenDatabaseSchema)
    conn <- DatabaseConnector::connect(dbProfileConnectionDetails)
    
    dbVisits <- DatabaseConnector::querySql(conn, renderedSql)

DatabaseConnector::disconnect(conn)

dbVisits <- dbVisits %>%
  mutate(,
    label = paste0(
      "Partner: ", OEN_ID, "<br>",
      "Visit Type: ", VISIT_ANCESTOR_CONCEPT_NAME, "<br>",
      "Proportion Persons: ", percent(PROP, accuracy = 0.1)
    )
  )

# Build the ggplot
visit_plot <- ggplot(dbVisits, aes(x = VISIT_ANCESTOR_CONCEPT_NAME, y = OEN_ID, fill = PROP, text = label)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(
    name = "Proportion of persons",
    guide = guide_colorbar(barwidth = 10, barheight = 1)
  ) +
  labs(
    title = "Proportion of Persons by Data Partner and Visit Type",
    x = "Visit Concept",
    y = "Data Partner (OEN_ID)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "right"
  )

# Convert to interactive plot with custom tooltip
plotly::ggplotly(visit_plot, tooltip = "text")

```

### Domain Coverage

```{r echo=FALSE} 

sql <-  "select oen_id, 'prop_persons_w_condition' as domain, stratum_2
from @oen_database_schema.db_profiles dp 
where analysis_id = 2004
 and stratum_1 = '1000000'
union
select oen_id, 'prop_persons_w_drugs' as domain, stratum_2
from @oen_database_schema.db_profiles dp 
where analysis_id = 2004
 and stratum_1 in ('0100000','100000')
union
select oen_id, 'prop_persons_w_devices' as domain, stratum_2
from @oen_database_schema.db_profiles dp 
where analysis_id = 2004
 and stratum_1 in ('0010000','10000')
union
select oen_id, 'prop_persons_w_measurements' as domain, stratum_2
from @oen_database_schema.db_profiles dp 
where analysis_id = 2004
 and stratum_1 in ('0001000','1000')
union
select oen_id, 'prop_persons_w_death' as domain, stratum_2
from @oen_database_schema.db_profiles dp 
where analysis_id = 2004
 and stratum_1 in ('0000100','100')
union
select oen_id, 'prop_persons_w_procedures' as domain, stratum_2
from @oen_database_schema.db_profiles dp 
where analysis_id = 2004
 and stratum_1 in ('0000010','10')
union
select oen_id, 'prop_persons_w_observations' as domain, stratum_2
from @oen_database_schema.db_profiles dp 
where analysis_id = 2004
 and stratum_1 in ('0000001','1')
 "
    
    renderedSql <- SqlRender::render(sql, oen_database_schema = oenDatabaseSchema)
    conn <- DatabaseConnector::connect(dbProfileConnectionDetails)
    
    dbDomains <- DatabaseConnector::querySql(conn, renderedSql)

DatabaseConnector::disconnect(conn)

# Create heatmap with ggplot
dbDomains <- dbDomains %>%
  mutate(
    STRATUM_2 = as.numeric(STRATUM_2),
    DOMAIN = gsub("prop_persons_w_", "", DOMAIN),
    DOMAIN = gsub("_", " ", DOMAIN),
    DOMAIN = tools::toTitleCase(DOMAIN),
    tooltip_text = paste0(
      "Data Partner: ", OEN_ID, "<br>",
      "Domain: ", DOMAIN, "<br>",
      "Coverage: ", percent(STRATUM_2, accuracy = 0.1)
    )
  )

# Create ggplot object with tooltip text
domain_plot <- ggplot(dbDomains, aes(x = DOMAIN, y = OEN_ID, fill = STRATUM_2, text = tooltip_text)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(
    limits = c(0, 1),
    labels = label_percent(accuracy = 1),
    name = "Coverage (%)"
  ) +
  labs(
    title = "Proportion of Persons with a Record in Each OMOP Domain",
    x = "OMOP Domain",
    y = "Data Partner (OEN_ID)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    axis.text.y = element_text(size = 6)
  )

# Convert to interactive plotly chart with formatted tooltip
ggplotly(domain_plot, tooltip = "text")
```
